CODECHECK: An open-science initiative to facilitate sharing of computer programs and results presented in scientific publications.

### 2020-12-14 codecheck-paper.tex
add further reference, try to fix long URL in footnote, fix URLs

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage{url}
\usepackage{setspace}

@@

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\item Should we make code and data sharing a prerequisite or leave that door open?
\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
\item Should we ask the people currently listed in the acknowledgements?
\item revisit \url{https://github.com/codecheckers/discussion}
\end{itemize}

@@

Yet, CODECHECK underlies the same limitations as peer review
in general and is closely connected to larger disruptions and challenges
in scholarly communication
\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, 
including the tensions between commercial publishing and reviewer's often
free labour, and a global pandemic jumbling up scientific publishing
and exposing a broader general audience to science, even preliminary 
results \cite{munafo_what_2020}.
Even more, it is clear that establishing CODECHECK-like systems can be 
impeded by and must be seen in connection to much larger issues in 
science, such as broken metrics or malpractices triggered by publication
pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
We certainly do not want the binary attribute of "CODECHECK completed"
to become a factor in bibliometric approaches for performance assesments.
% Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745
While developed for the current "paper"-centric publication process and
being compatible with different styles of peer review, the CODECHECK 
principles would also work well with novel publication paradigms, e.g.,
peer-reviewed computational notebooks \cite{earthcube_new_2020},
iterative and granular communication of scientific outputs,
such as Octopus\footnote{\url{https://science-octopus.org/}}, 
articles with live-code \cite{perkel_pioneering_2019-1}
\footnote{For example, eLife's ERA, see 
\url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
and completely new visions for scholarly communication and peer review
\footnote{Cf., for example, \emph{A modern vision for peer review} by 
Amy J. Ko: \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
An explicit segmentation of research steps could even make the focus 
of a CODECHECK easier by only checking the ``analysis'' sub-publication.
The discovery of CODECHECKs could be increased by depositing the reports
into public databases of reproductions, such as \emph{SciGen.Report}.
Public researcher profiles, such as ORCID, may consider different
types of reviewer activity to capture the contributions of independent 
code execution to science.
Notably, the discussed limitations are largely self-imposed for easier
acceptance and evolutionary integration, as to not break the current 
system and increase demands over time without leaving practitioners behind.

@@

With the implementation of CODECHECK workflows, the question of training
codecheckers will become relevant. We expect both a mentoring scheme within
the CODECHECK community (experienced codecheckers provide on the job training
or serve as fallback advisors) as well as collaboration with reproducible
research initiatives such as
ReproHack\footnote{https://reprohack.github.io/reprohack-hq/},
ReproducibiliTea\footnote{https://reproducibilitea.org/}
\cite{fitzgibbon_brewing_2020},
and Repro4Everyone\footnote{https://repro4everyone.org/}
\cite{auer_reproducibility_2020},
to be able to grow the pool of codecheckers as needed.
The initial reaction of the scientific community, where up to today over 
twenty people spontaneously signed up as volunteer codecheckers after talks
on the initiative, makes us feel optimistic that the time is right for 
putting scholarly peer review on the path to facilitate sharing and execution
of computer programs and documenting reproducibility of results.

>>>>

\documentclass[12pt]{article}
\usepackage{f1000_styles}
\usepackage[numbers]{natbib}
\usepackage{hyperref}
\usepackage[hyphenbreaks]{breakurl}
\usepackage[hyphens]{url}
\usepackage{setspace}

@@

\begin{itemize}
\item Check with Giuliano re: ``code works''. Write to Naomi.
\item Use "execution" insteand of "sharing" in title?
\item Can we introduce "research compendium" as a term?
\item Should we make code and data sharing a prerequisite or leave that door open?
\item Take a look at the Lancet's statistical review process (areas for improvement by authors) \url{https://www.thelancet.com/journals/lancet/article/PII0140-6736(92)90409-V/fulltext}
\item How to relate to "reproducibility does not matter but trust does", see \url{https://gmbecker.github.io/MayInstituteKeynote2019/outline.html}
\item Should we ask the people currently listed in the acknowledgements?
\item revisit \url{https://github.com/codecheckers/discussion}
\item 
\end{itemize}

@@

Yet, CODECHECK underlies the same limitations as peer review
in general and is closely connected to larger disruptions and challenges
in scholarly communication
\cite{eglen_recent_2018,tennant_ten_2019,fyfe_mission_2019}, 
including the tensions between commercial publishing and reviewer's often
free labour, and a global pandemic jumbling up scientific publishing
and exposing a broader general audience to science, even preliminary 
results \cite{munafo_what_2020}.
Even more, it is clear that establishing CODECHECK-like systems can be 
impeded by and must be seen in connection to much larger issues in 
science, such as broken metrics or malpractices triggered by publication
pressure \cite{piwowar_altmetrics:_2013,nosek_promoting_2015}.
We certainly do not want the binary attribute of "CODECHECK completed"
to become a factor in bibliometric approaches for performance assesments.
% Mittermaier, Peer Review and Bibliometrics: http://hdl.handle.net/2128/22745
While developed for the current "paper"-centric publication process and
being compatible with different styles of peer review, the CODECHECK 
principles would also work well with novel publication paradigms, e.g.,
peer-reviewed computational notebooks \cite{earthcube_new_2020},
iterative and granular communication of scientific outputs,
such as Octopus\footnote{\url{https://science-octopus.org/}}, 
articles with live-code \cite{perkel_pioneering_2019-1}
\footnote{For example, eLife's ERA, see 
\url{https://elifesciences.org/labs/dc5acbde/welcome-to-a-new-era-of-reproducible-publishing}.},
decentralized infrastructure and public reviewer reputation systems
\cite{tenorio-fornes_towards_2019},
and completely new visions for scholarly communication and peer review
\footnote{Cf., for example, \emph{A modern vision for peer review} by 
Amy J. Ko: \url{https://medium.com/bits-and-behavior/a-modern-vision-for-peer-review-d5f73f0fae07}}.
An explicit segmentation of research steps could even make the focus 
of a CODECHECK easier by only checking the ``analysis'' sub-publication.
The discovery of CODECHECKs could be increased by depositing the reports
into public databases of reproductions, such as \emph{SciGen.Report}.
Public researcher profiles, such as ORCID, may consider different
types of reviewer activity to capture the contributions of independent 
code execution to science.
Notably, the discussed limitations are largely self-imposed for easier
acceptance and evolutionary integration, as to not break the current 
system and increase demands over time without leaving practitioners behind.

@@

With the implementation of CODECHECK workflows, the question of training
codecheckers will become relevant. We expect both a mentoring scheme within
the CODECHECK community (experienced codecheckers provide on the job training
or serve as fallback advisors) as well as collaboration with reproducible
research initiatives such as
ReproHack\footnote{\url{https://reprohack.github.io/reprohack-hq/}},
ReproducibiliTea\footnote{\url{https://reproducibilitea.org/}}
\cite{fitzgibbon_brewing_2020},
and Repro4Everyone\footnote{\url{https://repro4everyone.org/}}
\cite{auer_reproducibility_2020},
to be able to grow the pool of codecheckers as needed.
The initial reaction of the scientific community, where up to today over 
twenty people spontaneously signed up as volunteer codecheckers after talks
on the initiative, makes us feel optimistic that the time is right for 
putting scholarly peer review on the path to facilitate sharing and execution
of computer programs and documenting reproducibility of results.
